---
title: "COVID-19 Analysis"
author: "`r praise::praise('${Adjective} Actuary')`"
date: "`r lubridate::today()`"
output: 
    powerpoint_presentation:
    word_document:
        toc: true
params:
    states: "All"
editor_options: 
  chunk_output_type: console
---

> Data from the COVID-19 Data Repository by the Center for Systems Science and 
> Engineering (CSSE) at Johns Hopkins University. See https://github.com/CSSEGISandData/COVID-19 
> for the raw data and additional information on sources. 
> Click [here](https://www.arcgis.com/apps/opsdashboard/index.html#/bda7594740fd40299423467b48e9ecf6)
> to access the JHU CSSE COVID-19 Visual Dashboard.


*This analysis applies to `r if(params$states[[1]] == "All") "all states" else glue::glue_collapse(params$states, sep = ", ", last = ", and ")`.*

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.height = 6, fig.width = 8)

library(tidyverse)
library(lubridate)
library(gghighlight)

# read in data
# if you don't have a copy of this file, run the script "download_cleanup.R"
dat <- readRDS("covid19-us.rds")

if (params$states[[1]] != "All") dat <- filter(dat, State %in% params$states)

# color theme
soa_colors <- rgb(
    c(  2, 119, 210, 253, 186, 226),
    c( 77, 196,  49, 206, 191, 127),
    c(124, 213,  56,   7,  51,  38),
    maxColorValue = 255
)


# plotting theme
theme_set(theme_light() + 
              theme(strip.background = element_rect(fill = soa_colors[[1]]))
)


scale_color_soa <- function (...) {
    scale_color_manual(values = soa_colors, ...)
}

prep <- function(dat) {
    dat %>%
        group_by(cal_date, .add = TRUE) %>%
        summarize(Cases = sum(tot_cases),
                  Deaths = sum(tot_deaths),
                  `Crude Mortality` = sum(tot_deaths) / sum(tot_cases),
                  `Crude Infection Rate` = sum(tot_cases) / sum(Population),
                  .groups = "drop") %>%
        pivot_longer(c("Cases", "Deaths", 
                       "Crude Mortality", "Crude Infection Rate"), 
                     names_to = "metric", 
                     values_to = "value") %>% 
        drop_na(value)
}

```

```{r highlevel-stats}
total_infections <- sum(dat$cases)
total_deaths <- sum(dat$deaths)
total_population <- dat %>% filter(cal_date == max(cal_date)) %>% 
  pull(Population) %>% sum()
infection_rate <- total_infections / total_population
fatality_rate <- total_deaths / total_infections
```

## Summary Stats:

**Total infections**: `r scales::comma(total_infections)`

**Total deaths**: `r scales::comma(total_deaths)`

**Infection rate**: `r scales::percent(infection_rate, accuracy = 0.1)`

**Case fatality rate**: `r scales::percent(fatality_rate, accuracy = 0.1)`


# USA Cases and Deaths {.tabset}

## Totals

```{r counts-all}

dat_all <- prep(dat)

dat_all %>%
    filter(metric %in% c("Cases", "Deaths")) %>% 
    ggplot(aes(cal_date, value)) +
    geom_line(lwd = 1.05) +
    scale_y_continuous(labels = scales::comma) +
    facet_wrap(~ metric, scales = "free_y") + 
    labs(x = "Date", y = "Count")

```

## Crude Infection and Mortality

```{r rates-all}

dat_all %>%
    filter(str_detect(metric, "^Crude")) %>% 
    ggplot(aes(cal_date, value)) +
    geom_line(lwd = 0.9) +
    scale_y_continuous(labels = scales::percent) +
    facet_wrap(~ metric, scales = "free_y") + 
    labs(x = "Date", y = "Cumulative Rate")

```



# State Cases and Deaths {.tabset}

## Totals

```{r counts-states}

dat_state <- dat %>% group_by(State) %>% prep()

dat_state %>%
    filter(metric %in% c("Cases", "Deaths")) %>% 
    ggplot(aes(cal_date, value, color = State, group = State)) +
    geom_line(size = 1.1) +
    gghighlight(max(value), max_highlight = 6, 
                calculate_per_facet = TRUE, 
                unhighlighted_params = list(size = 0.9, alpha = 0.5)) + 
    scale_y_continuous(labels = scales::comma) +
    facet_wrap(~ metric, scales = "free_y") + 
    labs(x = "Date", y = "Count") + 
    scale_color_soa()

```


## Crude Infection and Mortality

```{r rates-states}

dat_state %>%
    filter(str_detect(metric, "^Crude")) %>% 
    ggplot(aes(cal_date, value, color = State)) +
    geom_line(lwd = 1.1) +
    gghighlight(max(mean(value)), max_highlight = 6, 
                calculate_per_facet = TRUE, 
                unhighlighted_params = list(size = 0.9, alpha = 0.5)) + 
    scale_y_continuous(labels = scales::percent) +
    facet_wrap(~ metric, scales = "free_y") + 
    labs(x = "Date", y = "Cumulative Rate") + 
    scale_color_soa()

```
